---
title: "Homework1_Guanlong Fu"
author: "Guanlong"
date: "2018Äê2ÔÂ1ÈÕ"
output: html_document
abstract: |
   This is homework 1 to learn to use R markdown to output well formulated documents with texts, figures and results. The    aim of the exercise is to do a monte-carlo simulation to empirically calcualte the cdf of the standard normal            distribution and compare to the result to the **true** cdf to get a idea of the size of 
   the magnitude. 
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Mathematical equation

We what approximate the cdf of the standard normal distribution N(0,1), which is given as follows:

$$
\Phi(t) = \int_{-\infty}^{t} \frac{1}{\sqrt{2\pi}} e^{-y^2/2} ~ dy,
$$
by the following equation:
$$
\hat{\Phi}(t)= \frac{1}{n}\sum_{i=1}^{n}I(X{i} \leq t ),                                                               
$$
First we write a function named "empirical_cdf_sdnormal to calculate the above estimated empirical cdf of N(0,1)


```{r function }
empirical_cdf_sdnormal <- function (n,t)
{
  x <- rnorm(n, mean = 0, sd = 1)
  b = 0
  for (i in 1:n) {
    if ( x[i] <= t) {b=b+1}
  }
  empirical_cdf <- b/n
  
  return(empirical_cdf)
  
}


```

Then We code nexted loops to calculate the value of $\hat{\Phi}(t)$ for each of three value of $n=\{ 10^{2},10^{3},10^{4}\}$ and for the each of the nine values of t. Where the matrix "Phi_hat" is the calclated cdf for the combnation of $3\times9=27$ (n,t) , and repeated for 100 times for each (n,t) combnation. "bias" is the corresponding bias for each (n,t) combnation in each repitition.

Note here "Phi_hat" is a $3\times900$ matrix, each rown represent the estimates corresponding to a n value, and every 100 column stands for estimates corresponding to a t value. the matrix is too large to be shown completely, so we will first row of each matrix


```{r loop}
Phi_hat <- matrix(0,3,900)  # each row represents a n value, each 100 clo represents a t value
bias <- matrix(0,3,900) # bias matrix

n <- c(100,1000,10000)
t <- c(0.0,0.67,0.84,1.28,1.65,2.32,2.58,3.09,3.72)

for (i in 1:length(n)) {
  for (j in 1: length(t)) {
     for (k in (1:100)) {
       Phi_hat[i,100*(j-1)+k] <- empirical_cdf_sdnormal(n[i],t[j])  # 100*(j-1)+k, for the each 100 elements corresponding to the same value of t
       bias[i,100*(j-1)+k] <- Phi_hat[i,100*(j-1)+k]-pnorm(t[j])    # calculate the bias
       
       }
    
  } 
  
}
Phi_hat[1,]
bias[1,]
```
The boxplot is generated for all the t values 
```{r  echo=FALSE}
output_bias <- matrix(bias, 300,9, byrow = FALSE)
df<-data.frame(output_bias)
library(reshape2)
meltData <- melt(df)
library(ggplot2)
p <- ggplot(meltData, aes(factor(variable), value)) 
p + geom_boxplot() + facet_wrap(~variable, scale="free")


```
